
% ABSTRACT IN ENGLISH
\chapter*{Abstract} 
Recent progress in the digitalization of industrial processes led to a rise in studies on the three-dimensional bin packing problem.
The problem consists in packing a set of items in the minimum number of bins without any overlap.
When considering real-world settings, the addition of new practical constraints is required.
Previous studies in other fields related to container loading and pallet loading have shown that static stability of the bins is a crucial aspect to consider.
Most solutions in the literature address the problem of vertical support implicitly by generating dense layers of items that are then stacked to form a bin.
\\
In this thesis, we formulate a mixed-integer linear programming model for the three-dimensional bin packing problem with a discretized version of the vertical support constraint.
We then propose a constructive heuristic that fills bins without explicitly building layered solutions or using filler material, as is typically done in the literature.
In addition, we modify the two-dimensional Extreme Points algorithm to consider vertical support.
Moreover, we introduce a beam-search algorithm that evaluates different placements made by our constructive heuristic and filters out duplicate solutions.
We also provide a set of instances with items sampled from a population of real-world products.
Finally, we validate the results from our algorithm using different datasets, both from the literature and from a case study of mixed-case palletization.
\\
\\
\textbf{Keywords:} three dimensional bin packing, vertical support, static stability, mixed-case palletization